---
title: | 
  | Assignment 4: Collaborating Together
  | Introduction to Applied Data Science
  | 2022-2023
author: |
  | Maurits de Haan
  | m.c.p.dehaan@students.uu.nl
  | https://github.com/MCPdeHaan
date: 20th of June, 2023
urlcolor: purple
linkcolor: purple
output: 
  pdf_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE, error=TRUE)
```

## Assignment 4: Collaborating Together

### Part 1: Contributing to another student's Github repository

In this assignment, you will create a Github repository, containing this document and the .pdf output, which analyzes a dataset individually using some of the tools we have developed.

This time, make sure to not only put your name and student e-mail in your Rmarkdown header, but also your Github account, as I have done myself.

However, you will also pair up with a class mate and contribute to each others' Github repository. Each student is supposed to contribute to another student's work by writing a short interpretation of 1 or 2 sentences at the designated place (this place is marked with **designated place**) in the other student's assignment.

This interpretation will not be graded, but a Github shows the contributors to a certain repository. This way, we can see whether you have contributed to a repository of a class mate.

**Question 1.1**: Fill in the **github username** of the class mate to whose repository you have contributed.
The username of the class mate I did the project with is: Wojtek331

### Part 2: Analyzing various linear models

In this part, we will summarize a dataset and create a couple of customized tables. Then, we will compare a couple of linear models to each other, and see which linear model fits the data the best, and yields the most interesting results.

We will use a dataset called `GrowthSW` from the `AER` package. This is a dataset containing 65 observations on 6 variables and investigates the determinants of economic growth. First, we will try to summarize the data using the `modelsummary` package.

```{r, warning=FALSE, message=FALSE}
library(AER)
data(GrowthSW)
```

One of the variables in the dataset is `revolutions`, the number of revolutions, insurrections and coup d'etats in country $i$ from 1965 to 1995.

**Question 2.1**: Using the function `datasummary`, summarize the mean, median, sd, min, and max of the variables `growth`, and `rgdp60` between two groups: countries with `revolutions` equal to 0, and countries with more than 0 revolutions. Call this variable `treat`. Make sure to also write the resulting data set to memory. Hint: you can check some examples [here](https://vincentarelbundock.github.io/modelsummary/articles/datasummary.html#datasummary).

```{r}
library(modelsummary); library(tidyverse)

GrowthSW <- GrowthSW %>%
  mutate(treat = ifelse(revolutions == 0, "No revolution", "Revolution"))

summary <- datasummary(growth + rgdp60 ~ treat*(Mean + median + sd + min + max), data = GrowthSW)
summary
```

**Designated place**: type one or two sentences describing this table of a fellow student below. For example, comment on the mean and median growth of both groups. Then stage, commit and push it to their github repository.
If I look at the table of Wojtek, I see the that the value for GDP per capita in 1960 was 5283.32 on average for countries that had now revolutions and over two times smaller for countries that had a revolution (1988.67). This is logical, since the reason why a revolution occurred is often that something (often the economy) underperformed or the revolution had a big effect on the internal and external economy in a country, as shown by the lower average GDP per capita in 1960 for countries that had one or more revolutions. 


### Part 3: Make a table summarizing reressions using modelsummary and kable

In question 2, we have seen that growth rates differ markedly between countries that experienced at least one revolution/episode of political stability and countries that did not.

**Question 3.1**: Try to make this more precise this by performing a t-test on the variable growth according to the group variable you have created in the previous question.

```{r}
t.test(GrowthSW$growth ~ GrowthSW$treat)
```

**Question 3.2**: What is the $p$-value of the test, and what does that mean? Write down your answer below. The $p$-value that the t-test gives is **0.06871**. The goal of the p-value is to show what the probability is that we obtain the observed results, while assuming that the null hypothesis ($H_0$) is true. The general rule is: *The lower the p-value, the greater the statistical significance of the observed difference.* ([Alwan et. al., 2020](https://store.macmillanlearning.com/us/product/The-Practice-of-Statistics-for-Business-and-Economics/p/1319109004)). So, the p-value is greater than 0.05 which means that we cannot reject the null hypothesis ($H_0$) that there is no significant difference between "no revolution" and "revolution" in terms of the growth variable.

We can also control for other factors by including them in a linear model, for example:

$$
\text{growth}_i = \beta_0 + \beta_1 \cdot \text{treat}_i + \beta_2 \cdot \text{rgdp60}_i + \beta_3 \cdot \text{tradeshare}_i + \beta_4 \cdot \text{education}_i + \epsilon_i
$$

**Question 3.3**: What do you think the purpose of including the variable `rgdp60` is? Look at `?GrowthSW` to find out what the variables mean.

The purpose of including the variable `rgdp60` in the linear model is to control for the initial level of economic development. This is because the initial level of economic development, as shown by the GDP per capita in 1960, can have a significant impact on economic growth in the future, and this model wants to make sure that we are not just capturing the effect of the revolution on economic growth through the initial level of economic development.

The goal of this is, is to allows us to isolate the effect of the revolution on economic growth from the effect of the initial level of economic development. This is important because we want to know whether the revolution has a causal effect on economic growth, and we can only do this if we control for other factors that could also be affecting economic growth.

We now want to estimate a stepwise model. Stepwise means that we first estimate a univariate regression $\text{growth}_i = \beta_0 + \beta_1 \cdot \text{treat}_i + \epsilon_i$, and in each subsequent model, we add one control variable.

**Question 3.4**: Write four models, titled `model1`, `model2`, `model3`, `model4` (using the `lm` function) to memory. Hint: you can also use the `update` function to add variables to an already existing specification.

```{r}
library(AER)

# Model 1
model1 <- lm(growth ~ treat, data = GrowthSW)

# Model 2
model2 <- update(model1, ~. + rgdp60)

# Model 3
model3 <- update(model2, ~. + tradeshare)

# Model 4
model4 <- update(model3, ~. + education)
```

Now, we put the models in a list, and see what `modelsummary` gives us:

```{r}
list(model1, model2, model3, model4) |>
  modelsummary(stars=T,
  omit = c("AIC", "BIC", "Log-Likelihood", "Adj. R-squared", "Standard errors"),             
  keep = c("N", "R-squared")
) 
```

**Question 3.5**: Edit the code chunk above to remove many statistics from the table, but keep only the number of observations $N$, and the $R^2$ statistic.

**Question 3.6**: According to this analysis, what is the main driver of economic growth? Why?

The main driver of economic growth is `education`, since the coefficient on the variable `education` is positive and statistically significant in all four models. This suggests that an increase in education leads to an increase in economic growth. The other variables in the model also have some impact on economic growth, but the impact of education is the strongest. This is likely because education increases the skills of the workforce, which makes children more productive in the future.

Overall, the results of this analysis suggest that education is a key driver of economic growth. Policies that promote education are likely to lead to higher rates of economic growth in the long run.

**Question 3.7**: In the code chunk below, edit the table such that the cells (including standard errors) corresponding to the variable `treat` have a red background and white text. Make sure to load the `kableExtra` library beforehand.

```{r}
library(kableExtra)
library(modelsummary)

list(model1, model2, model3, model4) |>
  modelsummary(stars=T, gof_map = c("nobs", "r.squared")) %>%
  kable() %>%
  row_spec(background = "red", color = "white")
```

**Question 3.8**: Write a piece of code that exports this table (without the formatting) to a Word document.

```{r}

```

## The End
